### 2015-01-26 shevchenko/summaries.tex

\clearpage
\addcontentsline{toc}{chapter}{Summaries of all articles}
\markboth{}{}
\begingroup
\begin{center}
\Large \bfseries Summaries of all Articles
\end{center}

\selectlanguage{english}

\arthead{Full Name of author}{Article's name}

\begin{abstract}
Abstract.
\kwe keywords.
\end{abstract}

\selectlanguage{russian}

\endgroup