### 2015-02-01 shevchenko/summaries.tex

\clearpage
\addcontentsline{toc}{chapter}{Summaries of all articles}
\markboth{}{}
\begingroup
\begin{center}
\Large \bfseries Summaries of all Articles
\end{center}

\selectlanguage{english}

\arthead{Shevchenko~A.V., Tsybulin~I.V.}{An optimization algorithm for solving phase equilibrium problem}

\begin{abstract}
In the numerical simulation of multicomponent flows in many cases it is necessary to consider the possibility of phase transitions. In this paper we deal with the phase equilibrium problem for mixtures of several components, characteristic for hydrocarbon deposits (water, light and heavy oil). We assume that the phase equilibrium conditions are described in terms of the distribution coefficients (phase equilibrium constants). The phase equilibrium problem is to determine in what phases the mixture splits, their compositions and mole fraction, and the temperature, knowing  molar concentrations of all components, pressure and molar enthalpy of mixture. We construct an algorithm which consistently solve problems of this kind, based on the method of logarithmic barriers. The comparison of the developed algorithm with its ``standard'' analogue was made.

\kwe phase equilibrium, the method of logarithmic barriers, Gibbs potential, mathematical modeling.

\end{abstract}

\selectlanguage{russian}

\endgroup